---
title: "K-means clustering"
author: "Joshua Edefo"
date: "2024-07-27"
email: "edefojoshua2000@yahoo.com"
output: github_document
---
library
```{r setup, meessage =FALSE}
library(tidyverse)
```
Load the Iris dataset

```{r a}
data(USArrests)
head (USArrests)
str (USArrests)
```
put in data frame

```{r b}
df <- data.frame (USArrests)
```
Remove rows with missing values & Scale each variable

```{r c}
df <- na.omit(df)

df <- scale(df)

```
Methods of finding k

Elbow Plot: Next, create a plot of the number of clusters vs. the total within sum of squares (WCSS). 

The WCSS measures the variability within each cluster. We’ll use the fviz_nbclust() function from the factoextra package


```{r d}
library(cluster)
library(factoextra)
```
 Create plot of number of clusters vs. WCSS # elbow method

```{r e}
fviz_nbclust(df, kmeans, method = "wss")
# the optimal value is 4

```
Create plot of number of clusters vs. silhouette # silhuoette method

```{r f}
fviz_nbclust(df, kmeans, method = "silhouette")
# the optimal value is 2
```
Create plot of number of clusters vs. gap_stat # gap_stat method

```{r g}
fviz_nbclust(df, kmeans, method = "gap_stat")
# the optimal value is 2

```
 K- means Clustering when k = 4

```{r h}
# Specify the number of clusters (k)
k <- 2
# Perform k-means clustering
kmeans_result <- kmeans(df, centers = k, nstart = 25)
kmeans_result

```
K- means Clustering when k = 2

```{r i}
# Specify the number of clusters (k)
k <- 2
# Perform k-means clustering
kmeans_result <- kmeans(df, centers = k, nstart = 25)
kmeans_result


```
Add the cluster assignments to the original dataset

```{r j}
USArrests$Cluster <- as.factor(kmeans_result$cluster)
USArrests$Cluster

```
Visualize the clusters

```{r k}
library(ggplot2)

ggplot(USArrests, aes(x = Assault , y = UrbanPop, color = Cluster)) +
  geom_point() +
  labs(title = "K-means Clustering of USArrests", x = "Assault ", y = "UrbanPop")


```

session information

```{r l}
sessionInfo()

```
